---
title: "project"
author: "Hareesh"
date: "25 November 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(ROCR)
library(pROC)
library(caTools)
library(ROSE)
setwd("C:/Users/Administrator/Desktop/project")
```


```{r}
tr = read.csv("C:/Users/Administrator/Desktop/project/train.csv")
te = read.csv("C:/Users/Administrator/Desktop/project/test.csv")

te$is_promoted = NA  ## adding is_promoted column with NA values

df = rbind(tr , te) ## combined two datasets for EDA
```

## dim
```{r}
dim(df)  ## gives number of rows and columns
```


## structure of data
```{r}
str(df) ## shows the type of data
```

## summary of the data
```{r}
summary(df) ## gives the data distribution
```


## missing values
```{r}
colSums(is.na(df))
## previous_year_rating has missing values of 3705
```



```{r}
length(unique(df$employee_id))  ## number of unique records
```

## Univariate Analysis

## Department
```{r}

plot(df$department)
```

## region
```{r}
plot(df$region)
```
## education
```{r}
plot(df$education)

```
## gender
```{r}
plot(df$gender)
```

## recruitment channel
```{r}
plot(df$recruitment_channel)
```

## no.of trainings
```{r}

plot(as.factor(df$no_of_trainings))
```

## age distribution
```{r}
hist(df$age)
```

## previous year rating
```{r}

plot(as.factor(df$previous_year_rating))
```


## length of service

```{r}
plot(as.factor(df$length_of_service))
```


```{r}
plot(as.factor(df$awards_won.))
```

## is_promoted
```{r}
df$is_promoted = as.factor(df$is_promoted)
plot(df$is_promoted)
```

## bivariate analysis

## department vs is_promoted(==1)
```{r}

dept=tr%>%group_by(department)%>%summarise(freq = n())

tr %>%filter(is_promoted == 1)%>%group_by(department)%>%summarise(count = n())%>%arrange(desc(count))%>%left_join(dept) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion)) %>% ggplot(aes(x = reorder(department,-per_conversion),y = per_conversion,fill = department)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90),axis.title = element_blank(),axis.text.y = element_blank(),axis.ticks = element_blank()) + guides(fill = F) + geom_text(aes(label = paste0(per_conversion*100,"%")),position = position_stack(vjust = 0.5)) + labs(title = "Dept")


tr %>%filter(is_promoted == 1)%>%group_by(department)%>%summarise(count = n())%>%arrange(desc(count))%>%left_join(dept) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))
```


## region vs is_promoted(==1)
```{r}
region=tr%>%group_by(region)%>%summarise(freq = n())

tr %>%filter(is_promoted == 1)%>%group_by(region)%>%summarise(count = n())%>%left_join(region) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))%>%ggplot(aes(x = reorder(region,-per_conversion),y = per_conversion,fill = region)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90),axis.title = element_blank(),axis.text.y = element_blank(),axis.ticks = element_blank()) + guides(fill = F) + geom_text(aes(label = paste0(per_conversion*100,"%")),position = position_stack(vjust = 0.5)) + labs(title = "region")


tr %>%filter(is_promoted == 1)%>%group_by(region)%>%summarise(count = n())%>%left_join(region) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))
```


## education vs is_promoted(==1)
```{r}
education=tr%>%group_by(education)%>%summarise(freq = n())

levels(tr$education)[1] <- "others"

tr %>%filter(is_promoted == 1)%>%group_by(education)%>%summarise(count = n())%>%left_join(education) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))%>%ggplot(aes(x = reorder(education,-per_conversion),y = per_conversion,fill = education)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90),axis.title = element_blank(),axis.text.y = element_blank(),axis.ticks = element_blank()) + guides(fill = F) + geom_text(aes(label = paste0(per_conversion*100,"%")),position = position_stack(vjust = 0.5)) + labs(title = "education")


tr %>%filter(is_promoted == 1)%>%group_by(education)%>%summarise(count = n())%>%left_join(education) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))
```

## gender vs is_promoted(==1)

```{r}
gender=tr%>%group_by(gender)%>%summarise(freq = n())

tr %>%filter(is_promoted == 1)%>%group_by(gender)%>%summarise(count = n())%>%left_join(gender) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))%>%ggplot(aes(x = reorder(gender,-per_conversion),y = per_conversion,fill = gender)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90),axis.title = element_blank(),axis.text.y = element_blank(),axis.ticks = element_blank()) + guides(fill = F) + geom_text(aes(label = paste0(per_conversion*100,"%")),position = position_stack(vjust = 0.5)) + labs(title = "gender")


tr %>%filter(is_promoted == 1)%>%group_by(gender)%>%summarise(count = n())%>%left_join(gender) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))
```

## recruitment channel vs is_promoted(==1)

```{r}

recruitment_channel=tr%>%group_by(recruitment_channel)%>%summarise(freq = n())

tr %>%filter(is_promoted == 1)%>%group_by(recruitment_channel)%>%summarise(count = n())%>%left_join(recruitment_channel) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))%>%ggplot(aes(x = reorder(recruitment_channel,-per_conversion),y = per_conversion,fill = recruitment_channel)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90),axis.title = element_blank(),axis.text.y = element_blank(),axis.ticks = element_blank()) + guides(fill = F) + geom_text(aes(label = paste0(per_conversion*100,"%")),position = position_stack(vjust = 0.5)) + labs(title = "recruitment_channel")

tr %>%filter(is_promoted == 1)%>%group_by(recruitment_channel)%>%summarise(count = n())%>%left_join(recruitment_channel) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))
```


## no_oftrainings vs is_promoted(==1)

```{r}
no_of_trainings=tr%>%group_by(no_of_trainings)%>%summarise(freq = n())

tr %>%filter(is_promoted == 1)%>%group_by(no_of_trainings)%>%summarise(count = n())%>%left_join(no_of_trainings) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))%>%ggplot(aes(x = reorder(no_of_trainings,-per_conversion),y = per_conversion,fill = no_of_trainings)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90),axis.title = element_blank(),axis.text.y = element_blank(),axis.ticks = element_blank()) + guides(fill = F) + geom_text(aes(label = paste0(per_conversion*100,"%")),position = position_stack(vjust = 0.5)) + labs(title = "no_of_trainings")

tr %>%filter(is_promoted == 1)%>%group_by(no_of_trainings)%>%summarise(count = n())%>%left_join(no_of_trainings) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))
```

## awards won vs is_promoted(==1)

```{r}
awards_won.=tr%>%group_by(awards_won.)%>%summarise(freq = n())

tr %>%filter(is_promoted == 1)%>%group_by(awards_won.)%>%summarise(count = n())%>%left_join(awards_won.) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))%>%ggplot(aes(x = reorder(awards_won.,-per_conversion),y = per_conversion,fill = awards_won.)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90),axis.title = element_blank(),axis.text.y = element_blank(),axis.ticks = element_blank()) + guides(fill = F) + geom_text(aes(label = paste0(per_conversion*100,"%")),position = position_stack(vjust = 0.5)) + labs(title = "awards_won.")

tr %>%filter(is_promoted == 1)%>%group_by(awards_won.)%>%summarise(count = n())%>%left_join(awards_won.) %>%mutate(per_conversion = round(count/freq,2)) %>% arrange(desc(per_conversion))
```

## avg_training score vs is_promoted(==1)

```{r}
mean(tr$avg_training_score,na.rm = T)
tr %>%filter(is_promoted == 1) %>%summarise( Avg_score=mean(avg_training_score))


tr %>%filter(is_promoted == 1) %>%summarise( minimum_score=min(avg_training_score))
## mi9nimum score required to promoted
```

### missing value treatment
```{r}
# Education has 2290 missing values
#previous_year_rating has a missing values of 3705

levels(df$education)[1] <- "others"   ## imputed missing values with others


df$previous_year_rating[is.na(df$previous_year_rating)]=median(df$previous_year_rating,na.rm = T)
                                        ## impute missing value with median

df=df[-(which(is.na(df$avg_training_score))),]
                                      ## 1 missing value in avg training score and removed that row



```


```{r}
colSums(is.na(df))

```



### model building

## splitting in to train and test
```{r}

valid.df = df[is.na(df$is_promoted),]
dataset = df[!(is.na(df$is_promoted)), ]
data_part <- createDataPartition(dataset$is_promoted, 
                                 p = 0.7, list = F)
test <- dataset[-(data_part),] # 30% data goes here
train <- dataset[data_part,] # 70% here

table(train$is_promoted)
```


## logistic regression
```{r}
lg.model = glm((is_promoted) ~ .,data =train%>%select(-(employee_id)) , family = binomial)

summary(lg.model)

lg.predict = predict(lg.model , test%>%select(-c(is_promoted,employee_id)), type = "response")


lg.predict = if_else(lg.predict > 0.5 ,1,0)



```

```{r}
a <- factor(test$is_promoted)
b <- factor(lg.predict)
caret::confusionMatrix(b , a)


```




## roc curve
```{r}
rocplot <- function(pred, truth, ...) {
  predob = prediction(pred, truth)
  perf = performance(predob, "tpr", "fpr")
  plot(perf, ...)
  area <- auc(truth, pred)
  area <- format(round(area, 4), nsmall = 4)
  text(x=0.8, y=0.1, labels = paste("AUC =", area))

  # the reference x=y line
  segments(x0=0, y0=0, x1=1, y1=1, col="gray", lty=2)
}

rocplot(lg.predict,as.factor(test$is_promoted))
```



## decision tree
```{r}


decision.model = rpart(as.factor(is_promoted) ~ . ,data = train)


rpart.plot(decision.model , box.col = c("red","blue"))


pr = predict (decision.model, test )

tree.predict  = if_else(pr[,2] > pr[,1] , 1 , 0)
tree.predict = as.factor(tree.predict)

View(pr)
```


```{r}
confusionMatrix(tree.predict,a)
```


## roc curve and auc
```{r}

x= as.numeric(tree.predict)
y= as.numeric(test$is_promoted)
rocplot(x,y)
```


```{r}

rf.model = randomForest(as.factor(is_promoted) ~ .  , data = train)


rf.predict  = predict(rf.model , test , type = "prob")

rf.predict = as.data.frame(rf.predict)

rf.predict$pred = if_else(rf.predict$`0`  > rf.predict$`1` , 0 ,1)



caret::confusionMatrix(as.factor(rf.predict$pred),as.factor(test$is_promoted),positive ="1" )


```


```{r}
x= as.numeric(rf.predict$pred)
rocplot(x,y)
```


## logistic regreesion
```{r}
MLmetrics::F1_Score(test$is_promoted , lg.predict)

```
 

## random forest
```{r}
MLmetrics::F1_Score(test$is_promoted ,rf.predict$pred )
```


